---
title: "Class Activity"
output: 
  tufte::tufte_html:
    css: "lab.css"
    tufte_variant: "envisioned"
    highlight: pygments
link-citations: yes
---

**Instructions:** Work with a neighbor to answer the following questions, then we will discuss the activity as a class. To get started, download the [class activity template](https://sta279-f23.github.io/class_activities/ca_lecture_5_template.qmd) file.

## What will happen?

For each question, predict what will happen when the code is run. Then run the code and check whether your prediction was correct.

1. 

```{r, eval=F}
x <- 10

g02 <- function(x){
  x <- x + 1
  return(x)
}

g02(x)
x + 1
```

2.

```{r, eval=F}
x <- 10

g02 <- function(x){
  x <- x + 1
  return(x)
}

x <- g02(x)
x + 1
```

3. 

```{r, eval=F}
g02 <- function(y){
  y <- y - 1
  return(y)
}

g02(g02(20))
```

## Practice with anonymous functions

4. Use the `integrate()` and an anonymous function to find the area under the curve for the following functions:

* `y = x^2 - x` for x in $[0, 1]$
* `y = sin(x) + cos(x)` for x in $[-\pi, \pi]$
* `y = exp(x)/x` for x in $[10, 20]$

