---
title: "Exam 1 review"
output: 
  tufte::tufte_html:
    css: "lab.css"
    tufte_variant: "envisioned"
    highlight: pygments
link-citations: yes
---


## Practice with iteration in R

1. What will the following code return when you run it in R?

```{r, eval=F}
output <- rep(NA, 10)
for(i in 1:5){
  output[i] <- i
}

output[6]
```

2. What will the following code return when you run it in R?

```{r, eval=F}
output <- rep(0, 10)
for(i in 1:10){
  output[i] <- i
}

output[6]
```

3. What will the following code return when you run it in R?

```{r, eval=F}
output <- rep(0, 10)
for(i in 1:10){
  output[i] <- i
}

output[11]
```

4. What will the following code return when you run it in R?

```{r, eval=F}
output <- rep(1, 10)
for(i in 2:10){
  output[i] <- i + output[i-1]
}

output[5]
```

5. What will the following code return when you run it in R?

```{r, eval=F}
output <- rep(1, 10)
for(i in 2:10){
  output[i] <- i + output[i+1]
}

output[5]
```


## Practice with function scoping

6. What will be the output of the following code?

```{r}
x <- 10
test_fun <- function()
```

