---
title: "Final exam review"
output: 
  tufte::tufte_html:
    css: "lab.css"
    tufte_variant: "envisioned"
    highlight: pygments
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=F)
knitr::opts_chunk$set(warning=F)
```

Below are questions to help you study for the final exam. These are examples of the kinds of questions I might ask.

* This is not a practice exam. 
* The questions cover most, but not all, possible material for the exam.


# Writing code


## Practice with data wrangling

In each of the questions below, write code (in R or Python) to produce the output from the original data.

1. Original data:

```{r, echo=F, message=F, warning=F}
library(tidyverse)
library(palmerpenguins)

sub_data <- penguins |>
  slice_sample(n=10) |>
  select(species, island, bill_length_mm, bill_depth_mm)

sub_data
```

Output:

```{r, echo=F}
sub_data |>
  count(species, island)
```

2. Original data:

```{r, echo=F}
sub_data
```

Output:

```{r, echo=F}
sub_data |>
  group_by(island, species) |>
  summarize(mean_length = mean(bill_length_mm, na.rm=T))
```


3. Original data:

```{r, echo=F}
sub_data
```

Output:

```{r, echo=F}
sub_data |>
  mutate(bill_ratio = bill_length_mm/bill_depth_mm)
```


4. Original data:

```{r, echo=F}
sub_data
```

Output:

```{r, echo=F}
sub_data |>
  filter(species == "Adelie", 
         island == "Dream")
```

5. Do this question without explicitly listing the columns.

Original data:

```{r, echo=F}
ex_df <- data.frame(
  x1 = c(1, 2, 3),
  x2 = c("a", "b", "c"),
  x3 = c(3, 1, 4),
  y1 = c("d", "e", "f"),
  y2 = c(2, 7, 9),
  y3 = c(7, 1, 2)
)

ex_df
```

Output:

```{r, echo=F}
ex_df |>
  select(where(is.numeric))
```



6. Do this question without explicitly listing the columns.

Original data:

```{r, echo=F}
ex_df
```

Output:

```{r, echo=F}
ex_df |>
  select(starts_with("x"))
```


7. Do this question without explicitly listing the columns.

Original data:

```{r, echo=F}
ex_df
```

Output:

```{r, echo=F}
ex_df |>
  summarize(across(starts_with("x") & where(is.numeric), mean, .names = "mean_{.col}"))
```
8. Original data:

```{r, echo=F}
ex_df <- data.frame(
  id = c(1, 2, 3),
  x_1 = c(3, 1, 4),
  x_2 = c(5, 8, 9),
  y_1 = c(0, 1, 2),
  y_2 = c(2, 7, 9)
)

ex_df
```

Output:

```{r, echo=F}
ex_df |>
  pivot_longer(cols = -id, names_to = c("group", "obs"), names_sep = "_")
```

9. Original data:

```{r, echo=F}
ex_df <- data.frame(
  id = c(1, 1, 2, 2, 3, 3),
  group = c("x", "y", "x", "y", "x", "y"),
  value = sample(1:6, replace=T)
)

ex_df
```

Output:

```{r, echo=F}
ex_df |>
  pivot_wider(id_cols = id, names_from = group, values_from = value)
```


## Joins

In each of the following questions, write code to produce the desired output from the two input datasets. The code may involve additional wrangling steps, beyond a join.

10.

```{r, echo=F}
df1 <- data.frame(
  id = c(1, 2, 3),
  x = c(7, 9, 13)
)

df2 <- data.frame(
  id = c(1, 2, 4),
  y = c(10, 12, 14)
)
```

```{r}
df1

df2
```


Output:

```{r, echo=F}
df1 |>
  left_join(df2, join_by(id))
```


11.

```{r}
df1

df2
```


Output:

```{r, echo=F}
df1 |>
  inner_join(df2, join_by(id))
```

12.

```{r, echo=F}
df1 = data.frame(
  a_x = 1,
  a_y = 2,
  b_x = 2,
  b_y = 3
)

df2 = data.frame(
  id = c("a", "b"),
  z = c(4, 5)
)
```

```{r}
df1

df2
```


```{r, echo=F}
df1 |> 
  pivot_longer(cols = -c(), names_to = c("id", ".value"), names_sep = "_") |>
  left_join(df2, join_by(id))
```


## Regular expressions

Consider the following strings:

```{r, echo=F}
strings <- c(
  "George Washington: February 22, 1732",
  "Thomas Jefferson: April 13, 1743",
  "Abraham Lincoln: February 12, 1809",
  "Theodore Roosevelt: October 27, 1858"
)

strings
```

For each question below, fill in the R code to produce the desired output.

13.

```{r, eval=F}
str_extract(strings, ...)
```

```{r, echo=F}
str_extract(strings, ".+(?=:)")
```

14.


```{r, eval=F}
str_extract(strings, ...)
```

```{r, echo=F}
str_extract(strings, "(?<=: ).+")
```

15.

```{r, eval=F}
str_extract(strings, ...)
```

```{r, echo=F}
str_extract(strings, "(?<=\\s).+(?=:)")
```

16.

```{r, eval=F}
str_extract(strings, ...)
```

```{r, echo=F}
str_extract(strings, "(?<=, ).+")
```

## More regular expressions

Consider the following strings:

```{r, echo=F}
strings <- c("apple", "banana", "canteloupe", "durian",
             "eggplant", "french fries", "goat cheese",
             "pizza", "99 red balloons", "101 dalmatians",
             "route 66")
```

```{r}
strings
```


For each question below, fill in the R code to produce the desired output.

17.

```{r, eval=F}
str_subset(strings, ...)
```

```{r, echo=F}
str_subset(strings, "\\d+")
```

18.

```{r, eval=F}
str_subset(strings, ...)
```

```{r, echo=F}
str_subset(strings, "^\\d+")
```

19.

```{r, eval=F}
str_subset(strings, ...)
```

```{r, echo=F}
str_subset(strings, "a")
```

20.

```{r, eval=F}
str_subset(strings, ...)
```

```{r, echo=F}
str_subset(strings, "\\s")
```

21.

```{r, eval=F}
str_subset(strings, ...)
```

```{r, echo=F}
str_subset(strings, "(.)\\1{1}")
```



# Reading code

For each of the following questions, either write the output of the code, or explain why it gives an error. (Some questions will run correctly, others will cause errors)

22.

```{r, eval=F}
x <- list()
for(i in 1:10){
  x[i] <- i
}
x[2] + 1
```

23.

```{r, eval=F}
nsim <- 1000 # number of games
results <- rep(NA, nsim)

for(i in 1:nsim){
  # each game starts with the marker in the middle
  marker <- 0
  
  while(abs(marker) < 0){
    robotA <- runif(1, 0, 0.5)
    robotB <- runif(1, 0, 0.5)
    marker <- marker + robotA - robotB
  }
  
  # check whether robot A wins
  results[i] <- marker >= 0.5
}

# fraction of the time that robot A wins
mean(results)
```


24.

```{r, eval=F}
mat <- matrix(0, nrow=5, ncol=3)
for(i in 1:5){
  for(j in 1:3){
    mat <- i + j
  }
}

mat
```

25.

```{r, eval=F}
mat <- matrix(0, nrow=5, ncol=3)
for(i in 1:5){
  for(j in 1:3){
    mat[j, i] <- i + j
  }
}

mat
```

26.

```{r, eval=F}
mat <- matrix(1, 3, 3)
for(i in 2:3){
  for(j in 2:3){
    mat[i,j] <- mat[i-1, j-1] + mat[i, j-1]
  }
}

mat
```


27.

```{r, eval=F}
f1 <- function(x = 1){
  return(x + 1)
}
g1 <- function(x){
  return(f1() + x)
}

f1(g1(3))
```

28.

```{r, eval=F}
f1 <- function(n, groups){
  x <- matrix(1, nrow=n, ncol=n)
  unique_groups = unique(groups)
  means <- matrix(nrow = length(unique_groups), ncol = n)
  for(i in 1:length(unique_groups)){
    means[i,] <- colMeans(x[groups == unique_groups[i],])
  }
  
  return(means)
}

f1(5, groups = c(1, 1, 2, 2, 2))
```



# Improving code efficiency

29. Re-write the following code to run as efficiently as you can.

```{r, eval=F}
x <- c()
for(i in 1:100){
  x <- c(x, runif(1))
}
```

# Probability simulation

You have 40 cards, with 4 different colors. Cards for each color are numbered 1--10. Two cards are picked at random (without replacement). What is the probability that the two cards chosen have different numbers and different colors?

30. Write a simulation to estimate the probability.

# Another probability simulation

An election is held between two candidates. Candidate A wins the election with $p$ votes, while candidate B loses with $q < p$ votes. Given these final vote counts, what is the probability that, when the votes are tallied, candidate $A$ has more votes than candidate $B$ throughout the count? (So candidate A has more votes after 1 vote has been counted, after 2 votes have been counted, etc.)

31. Below is code that attempts to estimate this probability, with $p = 20$ and $q = 10$. It is very wrong. Explain why.

```{r, eval=F}
p <- 20
q <- 10
nsim <- 1000
votes <- rep(c(0, 1), times = c(q, p))
results <- rep(NA, nsim)

for(i in 1:nsim){
  shuffled_votes <- sample(votes, p+q, replace=F)
  results[i] <- sum(shuffled_votes) > sum(1 - shuffled_votes)
}

mean(results)
```

32. Write code that correctly estimates the probability, given $p$ and $q$.


# Statistical simulation

In homework, you learned about the k-means clustering algorithm. Given a dataset and a number of clusters $k$, the k-means algorithm divides the data into $k$ groups.

Suppose that you have data $X_1,...,X_n$. Each observation $X_i$ comes from one of three groups, with means $\mu_1, \mu_2, \mu_3$. Let $G_i$ be the group for $X_i$, and suppose that $X_i \sim N(\mu_{G_i}, 1)$.

We observe $X_1,...,X_n$, but we don't get to see the actual groups $G_1,...,G_n$. Instead, we are going to try to estimate the group assignments using the k-means algorithm, with $k=3$. We want to estimate the probability that we correctly assign groups for all observations $X_i$, and explore how that probability changes with $\mu_1, \mu_2$, and $\mu_3$.

33. Design a simulation study to answer this question. You do not need to write code, but you must describe each of the ADEMP steps in enough detail that I could implement your simulation study.

